---
title: "Loudness by Genre"
output: html_notebook
---

An analysis of music by Genre to see if loudness varies

It was believed that online streaming platform have reduced loudness. But does this have the same effect accross all genres of music?


 There are 26 genres so it is a total of 232,725 tracks.


If we are right and loudness in the music production is bad for hearing 
but that music streaming services like spotify 

The most popular genres in the top 40 are now
Hip-hop, Pop, Rock, Electronic are the most popular genres in the top-40

https://www.economist.com/graphic-detail/2018/02/02/popular-music-is-more-collaborative-than-ever

The data is in

https://www.kaggle.com/zaheenhamidani/ultimate-spotify-tracks-db/download

Explain why this way of looking at loudness matters


First load some data. This is a version of the full dataset but missing lots of columns. How we construct it is below

```{r}
library(stringr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(readr)

df2=read.csv("smalldata.csv",encoding = "UTF-8" )

head(df2)
```


How these graphs are made

https://micahallen.org/2018/03/15/introducing-raincloud-plots/




```{r}

p3 <- ggplot(df2,aes(x=genre,y=loudness, fill = genre))+
  geom_flat_violin(position = position_nudge(x = .2, y = 0),adjust = 2)+
  geom_point(position = position_jitter(width = .15), size = .25)+
  ylab('Loudness')+xlab('Genre')+coord_flip()+theme_cowplot()+guides(fill = FALSE)+
  geom_boxplot(width = .1, guides = FALSE, outlier.shape = NA, alpha = 0.5) +
 #  geom_boxplot(aes(x = as.numeric(genre), y = loudness),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +#+0.25
 # geom_errorbar(data = summary_loud, aes(x = genre, y = Mean, ymin = Mean-ci, ymax = Mean+ci), position = position_nudge(.25), colour = "BLACK", width = 0.1, size = 0.8)+

  ggtitle('Loudness by Genre')+
theme(plot.title = element_text(hjust = 0.5))

ggsave("GenreR.png", height=20, width=20)

p3

```

```{r}
summary_loud<-as.data.frame(tapply(df2$loudness, df2$genre, summary))
```


```{r}

summary_loud<-as.data.frame(summary(df2))

```


```{r}
summary_loud
```

```{r}
df=read.csv("spotify.csv",encoding = "UTF-8" )

head(df)
```


```{r}
names(df)[1] <- "genre"
head(df)
```


```{r}

#library(lavaan)


df<-dplyr::select(df, -c('artist_name', 'key','mode','time_signature','track_name','track_id','popularity','danceability','acousticness','duration_ms','energy','instrumentalness','liveness','speechiness','tempo','valence')) 


```


```{r}
summary(df)
```


This gets the stats on each genre. Median ,mean loudness and such

```{r}
#summary<-
tapply(df$loudness, df$genre, summary)
```

$`Children's Music`
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-36.721 -14.094 -11.286 -11.642  -8.614   0.948 

$`Children’s Music`
data['genre'] = data['genre'].str.replace('’','\'')

df<-str_replace_all(df$genre, "’", "'")
summary(df)





```{r}
head(df)
```


filter out uncommon genres or ones I think are repetition

```{r}

df2 <-filter(df, grepl('Dance|Rock|Pop|Classical', genre))
head(df2)

```

```{r}
# Write CSV in R
write.csv(df2, file = "smalldata.csv", row.names=FALSE)
```



